Bayesian inference of relative fitness on high-throughput pooled competition assays

### 2023-12-06 doc/_logistic_sim.qmd
fix links to scripts

>>>>

### Simulation details

This section lists the parameters used to generate the synthetic data used
throughout the main text. All the code to reproduce every simulation and figure
in the paper can be found in [the GitHub
repository](https://github.com/mrazomej/bayesian_fitness/tree/master). Unless
stated otherwise, the simulations were initialized following these steps:

1. The initial number of cells were defined as:
- Ancestor lineage: `n_init` $\times$ `frac_anc`.
- Neutral lineages: $n_0^{(n)} \sim \text{Poisson} 
\left( 
    \texttt{n\_init} \times \frac{\texttt{frac\_neutral}}{\texttt{n\_neutral}} 
\right)$ 
- Mutant lineages: $n_0^{(m)} \sim \text{LogNormal} 
\left( 
    \mu = \log(\texttt{n\_init} \times \frac{\texttt{frac\_mut}}{\texttt{n\_mut}}),
    \sigma = 2 
\right)$

where

- `n_init` is the initial number of cells.
- `frac_anc` $\in [0, 1]$ is the initial fraction of the culture that is
ancestor.
- `frac_neutral` $\in [0, 1]$ is the initial fraction of the culture that is
barcoded neutral lineages.
- `frac_mut` $\in [0, 1]$ is the initial fraction of the culture that is
barcoded lineages whose relative fitness we want to infer--hereafter "mutat"
barcodes.
- `frac_anc` + `frac_neutral` + `frac_mut` = 1
- `n_neutral` is the number of unique neutral lineage barcodes.
- `n_mut` is the number of unique mutant barcodes.

2. The distribution of growth rates $\lambda$ were defined as
- Ancestor lineage \& neutral lineages: $\lambda_a =$ `λ_a`. See details below
for specific values for each simulation.
- Mutant lineages: $\lambda_m \sim 
\text{truncated}-\mathcal{N}(
    \mu =$ `λ_mut_mean`,
    $\sigma =$ `λ_bc_std`,
    $\text{truncation limits} =$ `λ_trunc`
$)$. See details below for specific values for each simulation.

#### Fig 2.

The following is a `README.md` file automatically generated when simulating the
data for Fig. 2 in the main text:

- File location [`link to
file`](https://github.com/mrazomej/bayesian_fitness/blob/master/code/processing/data001_logistic_250bc_01env_01rep/sim_data001.jl)
- Number of mutant barcodes
`n_mut = 225`
- Number of neutral barcodes
`n_neutral = 25`
- Ancestral strain growth rate
`λ_a = 1.0`
- Carrying capacity
`κ = 1.0e10`
- Number of generations
`n_gen = 8`
- Initial number of cells
`n_init = κ / (2^(n_gen)) = 3.90625e7`
- Initial fraction of culture that is ancestor
`frac_anc = 0.93`
- Initial fraction of culture that is neutrals
`frac_neutral = 0.02`
- Initial fraction of culture that is mutants
`frac_mut = 0.04999999999999995`
- Mutant fitness distribution mean
`λ_mut_mean = λ_a * 1.005 = 1.005`
- Mutant fitness distribution standard deviation to sample growth rates
`λ_mut_std = 0.1`
- Mutant fitness distribution truncation ranges for growth rates
`λ_trunc = [λ_a .* 0.9999, λ_a * 1.5] = [0.9999, 1.5]`
- Gaussian noise distribution standard deviation
`σ_lognormal = 0.3`

#### Fig 3.

The following is a `README.md` file automatically generated when simulating the
data for Fig. 3 in the main text:

- File location [`link to
file`](https://github.com/mrazomej/bayesian_fitness/blob/master/code/processing/data003_logistic_1000bc_03env_01rep/sim_multienv_data003.jl)
- Number of environments
`n_env = 3`
- Environment transitions
`env_idx = [1, 2, 3, 1, 2, 3]`
- Number of mutant barcodes
`n_mut = 900`
- Number of neutral barcodes
`n_neutral = 100`
- Ancestral strain growth rate on each environment
`λ_a = [1.0, 2.0, 0.5]`
- Carrying capacity
`κ = 1.0e10`
- Number of generations
`n_gen = 8`
- Initial number of cells
`n_init = κ / (2^(n_gen)) = 3.90625e7`
- Initial fraction of culture that is ancestor
`frac_anc = 0.93`
- Initial fraction of culture that is neutrals
`frac_neutral = 0.02`
- Initial fraction of culture that is mutants
`frac_mut = 0.04999999999999995`
- Mutant fitness distribution mean
`λ_mut_mean = λ_a * 1.005 = [1.005, 2.01, 0.5025]
- Mutant fitness distribution standard deviation to sample growth rates
`λ_mut_std = 0.1`
- Mutant fitness distribution truncation ranges for growth rates
`λ_trunc = [λ_a .* 0.9995, λ_a .* 1.5] = [[[0.995, 1.99, 0.4975]], [[1.5, 3.0, 0.75]]]`
- Gaussian noise distribution standard deviation
`σ_lognormal = 0.125`

The variable `env_idx` lists the order of the environment cycles used in the
simulation. In this case, transitions were of the form `1 -> 2`, `2 -> 3`, and
`3 -> 1`.

#### Fig 4.

The following is a `README.md` file automatically generated when simulating the
data for Fig. 4 in the main text:

- File location [`link to
file`](https://github.com/mrazomej/bayesian_fitness/blob/master/code/processing/data002_logistic_1000bc_01env_02rep/sim_hierarchical_data002.jl)
- Number of experimental replicates
`n_data = 2`
- Number of mutant barcodes
`n_mut = 900`
- Number of neutral barcodes
`n_neutral = 100`
- Ancestral strain growth rate
`λ_a = 1.0`
- Carrying capacity
`κ = 1.0e10`
- Number of generations
`n_gen = 8`
- Initial number of cells
`n_init = κ / (2^(n_gen)) = 3.90625e7`
- Initial fraction of culture that is ancestor
`frac_anc = 0.93`
- Initial fraction of culture that is neutrals
`frac_neutral = 0.02`
- Initial fraction of culture that is mutants
`frac_mut = 0.04999999999999995`
- Mutant fitness distribution mean
`λ_bc_mean = λ_a * 1.005 = 1.005`
- Mutant fitness distribution standard deviation to sample growth rates
`λ_bc_std = 0.1`
- Mutant fitness distribution truncation ranges for growth rates
`λ_trunc = [λ_a .* 0.9999, λ_a * 1.5] = [0.9999, 1.5]`
- Gaussian noise distribution standard deviation
`σ_lognormal = 0.3`
- Standard deviation for between-experiment fitness distribution variability
`σ_exp = 0.015`

The variable `σ_exp` defines the variance used to sample the variability between
simulated experimental replicates. In other words, given a "hyper-fitness" value
$\lambda$, for each experimental replicate, the growth rate was defined as
$\lambda_j \sim \mathcal{N}(\lambda,$ `σ_exp` $)$.

#### Fig 6.

The following is a `README.md` file automatically generated when simulating the
data for Fig. 6 in the main text:

- File location [`link to
file`](https://github.com/mrazomej/bayesian_fitness/blob/master/code/processing/data007_logistic_hierarchicalgenotype_1000bc_01env_01rep/sim_data007.jl)
- Number of mutant barcodes
`n_mut = 900`
- Number of neutral barcodes
`n_neutral = 100`
- Ancestral strain growth rate
`λ_a = 1.0`
- Carrying capacity
`κ = 1.0e10`
- Number of generations
`n_gen = 8`
- Initial number of cells
`n_init = κ / (2^(n_gen)) = 3.90625e7`
- Initial fraction of culture that is ancestor
`frac_anc = 0.93`
- Initial fraction of culture that is neutrals
`frac_neutral = 0.02`
- Initial fraction of culture that is mutants
`frac_mut = 0.04999999999999995`
- Mutant fitness distribution mean
`λ_bc_mean = λ_a * 1.005 = 1.005`
- Mutant fitness distribution standard deviation to sample growth rates
`λ_bc_std = 0.1`
- Mutant fitness distribution truncation ranges for growth rates
`λ_trunc = [λ_a .* 0.9999, λ_a * 1.5] = [0.9999, 1.5]`
- Gaussian noise distribution standard deviation
`σ_lognormal = 0.3`
- Number of unique genotypes
`n_geno = Int64(n_mut / 10) = 90`

The variable `n_geno` defines the number of unique genotypes in the simulation.
In this simulation, we aimed to have ten unique barcodes on average for each
genotype. The actual number of genotypes per barcode was sampled from a
multinomial distribution with probability `1 / n_geno` for each genotype.
